---
title: "Life Expectancy: regresion analysis"
output:
  html_document:
    code_folding: hide
date: 2021-11-20
categories: ["R"]
tags: ["R Markdown", "plot", "life expectancy"]
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(collapse = TRUE,warning = F, message = F)
```


From the visualization in last blog, we expect a linear trend of life expectancy over the time. Is our assumption right? In this blog, we will check this intuition statistically.  

we formulate a linear model as following:

$$\hat{\text{lifeExp}_i} = \beta_0 + \beta_1 * \text{year}_i$$

The estimates of this linear regression model is shown in table below:

```{r, results='asis'}
library(tidyverse)
library(stargazer)
data = read.csv("../../../raw_data/08_gap-every-five-years.tsv", sep = "\t")


ols = lm(lifeExp ~ year, data)
stargazer(ols, type ="html")




```



From the estimates we reject the null hypothesis of no relationship between year and life expectancy, on average, life expectancy increases by 0.3259 year every year around the world.

However, if we look at distribution of the residuals, groupped by continent, we find there is a relation between residuals of regression model and the continents.



```{r}
data$residulas = ols$residuals


data %>% ggplot(aes(y = residulas, color = continent)) + geom_boxplot() + 
  labs(title = "residuals by continents")
```

